---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# csranks

<!-- badges: start -->
[![R-CMD-check](https://github.com/danielwilhelm/R-CS-ranks/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/danielwilhelm/R-CS-ranks/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `R` package `csranks` implements confidence sets for ranks as in [Mogstad, Romano, Shaikh, and Wilhelm (2023)](http://dwilhelm.userweb.mwn.de/papers/cwp0323.pdf).

## Installation

You can install the development version of csranks from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("danielwilhelm/R-CS-ranks")
```

## Example: PISA Ranking

This example illustrates the computation of confidence sets for the ranking of countries according students' achievements in mathematics. The following dataset contains data from the 2018 Program for International Student Assessment (PISA) study by the Organization for Economic Cooperation and Development (OECD), which can be downloaded at https://www.oecd.org/pisa/data/. Students' achievement is measured by the average mathematics test scores for all 15-year-old students in the study.


First, load the package `csranks` and the data:

```{r}
library(csranks)
attach(pisa)
```

The dataframe `pisa` looks like this:

```{r}
head(pisa)
```

Consider ranking countries according to their students' achievement in mathematics. The scores in `math_score` are estimates of these achievements. The countries' ranks can be estimated using the function `irank()`:

```{r}
math_rank <- irank(math_score)
math_rank
```

These are only estimated ranks as they are computed from estimates of achievements and therefore estimation uncertainty in the estimates translates to estimation uncertainty in the ranks. The following subsections discuss two different confidence sets for assessing such estimation uncertainty in ranks: marginal and simultaneous confidence sets. Explanations and details about the differences of these two confidence sets can be found in Mogstad, Romano, Shaikh, and Wilhelm (2020). The complete R script containing this example, as well as R scripts for PISA rankings by reading and science scores, can be found in the subdirectory `examples/`.

### Marginal Confidence Sets

The marginal confidence sets for the ranks are computed as follows:

```{r}
math_cov_mat <- diag(math_se^2)
CS_marg <- csranks(math_score, math_cov_mat, coverage=0.95, simul=FALSE, R=1000, seed=101)
math_rankL_marg <- CS_marg$L
math_rankU_marg <- CS_marg$U
```

where `math_rankL_marg` and `math_rankL_marg` contain the lower and upper bounds of the confidence sets for the ranks. They can be plotted by

```{r}
grid::current.viewport()

plotmarg <- plot(CS_marg, popnames=jurisdiction, title="Ranking of OECD Countries by 2018 PISA Math Score", 
	subtitle="(with 95% marginal confidence sets)", colorbins=4)
plotmarg
```

You can then save the graph by

```{r, eval = FALSE}
ggplot2::ggsave("mathmarg.pdf", plot=plotmarg)
```

### Simultaneous Confidence Sets

The simultaneous confidence sets for the ranks are computed as follows:

```{r}
CS_simul <- csranks(math_score, math_cov_mat, coverage=0.95, simul=TRUE, R=1000, seed=101)
math_rankL_simul <- CS_simul$L
math_rankU_simul <- CS_simul$U
```

where `math_rankL_simul` and `math_rankL_simul` contain the lower and upper bounds of the confidence sets for the ranks. They can be plotted by

```{r}
grid::current.viewport()

plotsimul <- plot(CS_simul, popnames=jurisdiction, title="Ranking of OECD Countries by 2018 PISA Math Score", 
	subtitle="(with 95% simultaneous confidence sets)", colorbins=4)

plotsimul
```

You can then save the graph by

```{r, eval = FALSE}
ggplot2::ggsave("mathsimul.pdf", plot=plotsimul)
```


# Reference
[Mogstad, Romano, Shaikh, and Wilhelm (2023), "Inference for Ranks with Applications to Mobility across Neighborhoods and Academic Achievements across Countries", forthcoming at Review of Economic Studies](http://dwilhelm.userweb.mwn.de/papers/cwp0323.pdf)

